Supporting Information for Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-09-30 paper.tex
subtle changes
Added the paper as submitted July 30 to prep for diff

Added widetext / table* where needed

Tried to make all FC things listed as 'Franck Condon factors' and HT things listed as 'Herzberg-Teller integrals'

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{drsive_spectrum.png}
	\caption{Contributions of $A, B$ (\autoref{ABterms_DR}) to HDFG spectrum for a simple two harmonic well system.
		(a) Potential energy surfaces for a two-well system, (b) 1D HDFG spectrum for $\Delta = 0.5$, $\omega_1 = \omega_{g1, g0}$, (c, d) real and imaginary parts of terms plotted in (b), respectively.
		The spectra are referenced to $\max{(A+B)}$. 
		It is assumed that $\omega_2 = \omega_3$, and that all molecular parameters used to simulate the spectra are positive.
		The vibrational states on the $|g)$ and $|e)$ manifolds are spaced 1600 cm$^{-1}$ apart, with the vibronic states $\ket{e,\tilde{v}'}$ given linewidths of 300 cm$^{-1}$ and $\omega_{eg}$ = 30000 cm$^{-1}$.
		Dotted lines in (b) denote the $\tilde{v}'$ = 0, 1, 2 vibronic resonances. 
		The vibronic one and two photon absorption operators are scaled such that $\abs{B}/\abs{A} \sim$ 0.1, following Chung and Ziegler. \cite{Ziegler1988}
		The Franck-Condon and Herzberg-Teller contributions to each of these terms are found in \autoref{T:contributions}.}
	\label{fig:doubres_spec}
\end{figure*}

@@

\begin{table}[!htbp]
	\caption{\label{T:contributions} Franck-Condon and/or Herzberg-Teller contributions to $A$, $B_1$, $B_2$ terms between $\ket{g,0}$, $\ket{g,1}$ and vibronic states $\ket{e,\tilde{0}}$, $\ket{e,\tilde{1}}$, $\ket{e,\tilde{2}}$ for \autoref{Hamiltonian} with an offset $\Delta = 0.5$.
	The bra and ket labels are written according to their order in $A$, $B_1$, $B_2$ as written in \autoref{ABterms_DR}.
	The Franck-Condon and Herzberg-Teller integral products tabulated here are plotted in the Supporting Information (Figure S2) as a function of $\Delta$.}
	\begin{ruledtabular}
		\begin{tabular}{cccc}\label{fcht}
			$2\omega_2$ Electronic Resonance & $\ket{e,\tilde{0}}$ & $\ket{e,\tilde{1}}$ & $\ket{e,\tilde{2}}$\\
			\hline  
			$A$ 
			& $\langle 1 | \tilde{0} \rangle \langle \tilde{0} | 0 \rangle = 0.31 $  
			& $\langle 1 | \tilde{1} \rangle \langle \tilde{1} | 0 \rangle = -0.27 $ 
			& $\langle 1 | \tilde{2} \rangle \langle \tilde{2} | 0 \rangle = -0.04$\\

@@

			$B_2$ 
			& $\mel{1}{Q}{\tilde{0}} \langle \tilde{0} | 0 \rangle = 0.70$  
			& $\mel{1}{Q}{\tilde{1}} \langle \tilde{1} | 0 \rangle = -0.07$ 
			& $\mel{1}{Q}{\tilde{2}} \langle \tilde{2} | 0 \rangle = 0.06$\\
		\end{tabular}
	\end{ruledtabular}
\end{table}

@@

In the post-resonance region, the amplitude takes on an unexpected lineshape.
This arises due to interference between the real and imaginary parts of $\gamma$ (\autoref{fig:doubres_spec}c, d).
The abnormal lineshape arises due to the sign of the Franck-Condon and Herzberg-Teller integrals, as well as the molecular parameters that drive the output of the vibronic transitions.
Such effects have been seen elsewhere in spontaneous, surface enhanced hyper-Raman studies. \cite{Milojevich2011}
The main source of interference in the HDFG spectra arises from the sign of the molecular parameters. 
In the HDFG spectrum presented in \autoref{fig:doubres_spec}, the signs of all molecular parameters are taken to be positive. 
Inverting the sign of one parameter eliminates the interference and restores the expected, purely Lorentzian lineshape (Figure S4).
It is important to note that these spectra simulate a single element of the hyperpolarizability tensor, $\gamma_{ijkl}$.
The true HDFG response must be averaged over the molecular frame, or, the true, HDFG response is a linear combination of different $\gamma_{ijkl}$ terms.
Nevertheless, if the resultant linear combination results in destructive interference between the $A$ and $B$ terms, it could be possible to elucidate both the amplitude and sign of $\Delta$ from a HDFG experiment.

@@

We investigate the impact of inhomogeneity on spectral output by convolving the HDFG output hyperpolarizability $\gamma_{ijkl}(\xi)$ with a normalized Lorentzian distribution $P(\xi)$ of half width $\sigma$ as \cite{Desiderio1979} 
\begin{equation}\label{contour}
	\gamma_{ijkl} = \int_{-\infty}^\infty \mathrm{d}\xi \ P(\xi) \gamma_{ijkl}(\xi)
	= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)}
\end{equation} 
where $\eta_{ijkl}$ contains all values in \autoref{drgamma_notaylor} not dependent upon $\xi$.
While Gaussian distributions are commonly used as a phenomenological description of inhomogeneous broadening, Lorentzian distributions have successfully described inhomogeneous broadening in many coherent spectroscopies and yield easily interpretable closed-form solutions. \cite{Dick83_1, Carlson1990line, RN410, Yurs2012}

>>>>

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{drsive_spectrum.png}
	\caption{Contributions of $A, B$ (\autoref{ABterms_DR}) to HDFG spectrum for a simple two harmonic well system.
		(a) Potential energy surfaces for a two-well system, (b) 1D HDFG spectrum for $\Delta = 0.5$, $\omega_1 = \omega_{g1, g0}$, (c, d) real and imaginary parts of terms plotted in (b), respectively.
		The spectra are referenced to $\max{(A+B)}$. 
		It is assumed that $\omega_2 = \omega_3$, and that all molecular parameters used to simulate the spectra are positive.
		The vibrational states on the $|g)$ and $|e)$ manifolds are spaced 1600 cm$^{-1}$ apart, with the vibronic states $\ket{e,\tilde{v}'}$ given linewidths of 300 cm$^{-1}$ and $\omega_{eg}$ = 30000 cm$^{-1}$.
		Dotted lines in (b) denote the $\tilde{v}'$ = 0, 1, 2 vibronic resonances. 
		The vibronic one and two photon absorption operators are scaled such that $\abs{B}/\abs{A} \sim$ 0.1, following Chung and Ziegler. \cite{Ziegler1988}
		The Franck-Condon factor and Herzberg-Teller integral contributions to each of these terms are found in \autoref{T:contributions}.}
	\label{fig:doubres_spec}
\end{figure*}

@@

\begin{table*}[!htbp]
	\caption{\label{T:contributions} Franck-Condon and/or Herzberg-Teller contributions to $A$, $B_1$, $B_2$ terms between $\ket{g,0}$, $\ket{g,1}$ and vibronic states $\ket{e,\tilde{0}}$, $\ket{e,\tilde{1}}$, $\ket{e,\tilde{2}}$ for \autoref{Hamiltonian} with an offset $\Delta = 0.5$.
	The bra and ket labels are written according to their order in $A$, $B_1$, $B_2$ as written in \autoref{ABterms_DR}.
	The Franck-Condon factor and Herzberg-Teller integral products tabulated here are plotted in the Supporting Information (Figure S2) as a function of $\Delta$.}
	\begin{ruledtabular}
		\begin{tabular}{cccc}\label{fcht}
			$2\omega_2$ Electronic Resonance & $\ket{e,\tilde{0}}$ & $\ket{e,\tilde{1}}$ & $\ket{e,\tilde{2}}$\\
			\hline  
			$A$ 
			& $\langle 1 | \tilde{0} \rangle \langle \tilde{0} | 0 \rangle = 0.31 $  
			& $\langle 1 | \tilde{1} \rangle \langle \tilde{1} | 0 \rangle = -0.27 $ 
			& $\langle 1 | \tilde{2} \rangle \langle \tilde{2} | 0 \rangle = -0.04$\\

@@

			$B_2$ 
			& $\mel{1}{Q}{\tilde{0}} \langle \tilde{0} | 0 \rangle = 0.70$  
			& $\mel{1}{Q}{\tilde{1}} \langle \tilde{1} | 0 \rangle = -0.07$ 
			& $\mel{1}{Q}{\tilde{2}} \langle \tilde{2} | 0 \rangle = 0.06$\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}

@@

In the post-resonance region, the amplitude takes on an unexpected lineshape.
This arises due to interference between the real and imaginary parts of $\gamma$ (\autoref{fig:doubres_spec}c, d).
The abnormal lineshape arises due to the sign of the Franck-Condon factor and Herzberg-Teller integrals, as well as the molecular parameters that drive the output of the vibronic transitions.
Such effects have been seen elsewhere in spontaneous, surface enhanced hyper-Raman studies. \cite{Milojevich2011}
The main source of interference in the HDFG spectra arises from the sign of the molecular parameters. 
In the HDFG spectrum presented in \autoref{fig:doubres_spec}, the signs of all molecular parameters are taken to be positive. 
Inverting the sign of one parameter eliminates the interference and restores the expected, purely Lorentzian lineshape (Figure S4).
It is important to note that these spectra simulate a single element of the hyperpolarizability tensor, $\gamma_{ijkl}$.
The true HDFG response must be averaged over the molecular frame, or, the true, HDFG response is a linear combination of different $\gamma_{ijkl}$ terms.
Nevertheless, if the resultant linear combination results in destructive interference between the $A$ and $B$ terms, it could be possible to elucidate both the amplitude and sign of $\Delta$ from a HDFG experiment.

@@

We investigate the impact of inhomogeneity on spectral output by convolving the HDFG output hyperpolarizability $\gamma_{ijkl}(\xi)$ with a normalized Lorentzian distribution $P(\xi)$ of half width $\sigma$ as \cite{Desiderio1979} 
\begin{widetext}
\begin{equation}\label{contour}
	\gamma_{ijkl} = \int_{-\infty}^\infty \mathrm{d}\xi \ P(\xi) \gamma_{ijkl}(\xi)
	= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)}
\end{equation} 
\end{widetext}
where $\eta_{ijkl}$ contains all values in \autoref{drgamma_notaylor} not dependent upon $\xi$.
While Gaussian distributions are commonly used as a phenomenological description of inhomogeneous broadening, Lorentzian distributions have successfully described inhomogeneous broadening in many coherent spectroscopies and yield easily interpretable closed-form solutions. \cite{Dick83_1, Carlson1990line, RN410, Yurs2012}

### 2024-09-30 si.tex
subtle changes
Added the paper as submitted July 30 to prep for diff

Added widetext / table* where needed

Tried to make all FC things listed as 'Franck Condon factors' and HT things listed as 'Herzberg-Teller integrals'

For an $n$-quantum excited state vibration $| \tilde{n} \rangle$, and a $m$-quantum ground state vibration $\ket{m}$, the $k^{th}$-order vibrational coupling term is
\begin{equation}
	\begin{split}
		\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \ \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q \ N_n H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k N_m H_m(q) \exp(-\frac{q^2}{2}). \\
		&= N_n N_m \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ H_n(q-\Delta) q^k H_m(q) \exp(-q^2 + q\Delta)
	\end{split}\label{key}
\end{equation}
The integral corresponds to Franck-Condon factors for $k=0$ and Herzberg-Teller matrix elements for $k=1$.
The vibrational coupling integrals are written assuming the expansion of the dipole moment is taken about the equilibirum point of the ground state geometry ($q=0$), as is done in the main text.
Note that in the limit $\Delta \rightarrow 0$, typical harmonic oscillator selection rules are restored, as the two wells have the same vibrational Hamiltonian and thus share an identical subspace of orthonormal eigenstates.

@@

\newpage
\section{Impact of displacement, resonance frequencies, vibronic linewidth and molecular parameters on HDFG Spectra}
Inspection of the $A,B$ terms in the main text show that the HDFG output is dependent upon the relative size of products of Franck-Condon and Herzberg-Teller integrals.
For clarity, we plot the relevant product factors to assist in understanding the impact of offset on the HDFG spectra (\autoref{fig:fcht_product}).
These plots highlight the important distinction that $B$ term contributions to HDFG can dominate in the limit of small offset.
This is expected, as when $\Delta \rightarrow 0$, the overlap integral, \autoref{key}, returns typical harmonic oscillator selection rules.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/fchtproduct.png}
	\caption{Products of Franck-Condon and Herzberg-Teller integrals relevant to the $A, B_1, B_2$ terms in this manuscript as a function of offset, $\Delta$.} 
	\label{fig:fcht_product}
\end{figure}

In the main text, the potential well offset ($\Delta$) was taken as $\Delta = 0.5$.
Here, we vary $\Delta$ to investigate its impact on the structure of the $\abs{A+B}$ spectrum.
The dependence of relevant Franck-Condon factors and Herzberg-Teller factors on displacement are plotted in \autoref{fig:fcht} as a function of offset, $\Delta$.
\autoref{fig:fcht} shows how different displacements, vibrational frequencies, and vibronic linewidths will alter the vibronic spectrum.
The products of Franck-Condon and Herzberg-Teller integrals relevant to the $A,B$ coefficents discussed in the main text are plotted in \autoref{fig:fcht_product}.

@@

\section{Evaluating impact of static inhomogeneity on HDFG output}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated for $c=-1$ (electronic state and vibrational state energy fluctuations are anti-correlated) and $c=1$ (electronic state and vibrational state energy fluctuations are correlated).
Integrals such as those found in \autoref{generalcontour} can be evaluated by using the Residue theorem. %@dan i had too much white space, needed something to fill in... delete if you want.
If a complex valued function $f(z)$ is integrated along a positively oriented closed contour,%guessing on this one
\begin{subequations}\label{eq:generalcontour}
	\begin{equation}
		\oint \mathrm{d}z f(z) = 2\pi i \sum_k \Res[f(z_k)]
	\end{equation}
where the residue of an n$^{\text{th}}$ order pole at $z = z_k$, $\Res[f(z_k)]$ is evaluated as
	\begin{equation}
		\Res[f(z_k)] = \lim_{z \rightarrow z_k} \frac{1}{\left(n-1\right)!} \frac{\mathrm{d}^{n-1}}{\mathrm{d}z^{n-1}} \left[f(z) (z-z_k) \right]
	\end{equation}
For first order poles (the cases considered below), $n=1$, so that 
	\begin{equation}
		\oint \mathrm{d}z f(z) = 2\pi i \sum_k \lim_{z \rightarrow z_k} f(z) (z-z_k)
	\end{equation}
\end{subequations}
Note that in \autoref{eq:generalcontour}, the integrals are closed along a contour which wraps all the poles.

>>>>

For an $n$-quantum excited state vibration $| \tilde{n} \rangle$, and a $m$-quantum ground state vibration $\ket{m}$, the $k^{th}$-order vibrational coupling term is
\begin{equation}
	\begin{split}
		\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \ \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q \ N_n H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k N_m H_m(q) \exp(-\frac{q^2}{2}). \\
		&= N_n N_m \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ H_n(q-\Delta) q^k H_m(q) \exp(-q^2 + q\Delta)
	\end{split}\label{key}
\end{equation}
The integral corresponds to Franck-Condon factors for $k=0$ and Herzberg-Teller integrals for $k=1$.
The vibrational coupling integrals are written assuming the expansion of the dipole moment is taken about the equilibirum point of the ground state geometry ($q=0$), as is done in the main text.
Note that in the limit $\Delta \rightarrow 0$, typical harmonic oscillator selection rules are restored, as the two wells have the same vibrational Hamiltonian and thus share a subspace of orthonormal eigenstates.

@@

\newpage
\section{Impact of displacement, resonance frequencies, vibronic linewidth and molecular parameters on HDFG Spectra}
Inspection of the $A,B$ terms in the main text show that the HDFG output is dependent upon the relative size of products of Franck-Condon factors and Herzberg-Teller integrals.
For clarity, we plot the relevant product factors to assist in understanding the impact of offset on the HDFG spectra (\autoref{fig:fcht_product}).
These plots highlight the important distinction that $B$ term contributions to HDFG can dominate in the limit of small offset.
This is expected, as when $\Delta \rightarrow 0$, the overlap integral, \autoref{key}, returns typical harmonic oscillator selection rules.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/fchtproduct.png}
	\caption{Products of Franck-Condon factors and Herzberg-Teller integrals relevant to the $A, B_1, B_2$ terms in this manuscript as a function of offset, $\Delta$.} 
	\label{fig:fcht_product}
\end{figure}

In the main text, the potential well offset ($\Delta$) was taken as $\Delta = 0.5$.
Here, we vary $\Delta$ to investigate its impact on the structure of the $\abs{A+B}$ spectrum.
The dependence of relevant Franck-Condon factors and Herzberg-Teller integrals on displacement are plotted in \autoref{fig:fcht} as a function of offset, $\Delta$.
\autoref{fig:fcht} shows how different displacements, vibrational frequencies, and vibronic linewidths will alter the vibronic spectrum.
The products of Franck-Condon factors and Herzberg-Teller integrals relevant to the $A,B$ coefficents discussed in the main text are plotted in \autoref{fig:fcht_product}.

@@

\section{Evaluating impact of static inhomogeneity on HDFG output}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated for $c=-1$ (electronic state and vibrational state energy fluctuations are anti-correlated) and $c=1$ (electronic state and vibrational state energy fluctuations are correlated).
Integrals such as those found in \autoref{generalcontour} can be evaluated by using the Residue theorem. %@dan i had too much white space, needed something to fill in... delete if you want.
Recall that if a complex valued function $f(z)$ is integrated along a positively oriented closed contour,%guessing on this one
\begin{subequations}\label{eq:generalcontour}
	\begin{equation}
		\oint \mathrm{d}z f(z) = 2\pi i \sum_k \Res[f(z_k)]
	\end{equation}
where the residue of an n$^{\text{th}}$ order pole at $z = z_k$, $\Res[f(z_k)]$ is evaluated as
	\begin{equation}
		\Res[f(z_k)] = \lim_{z \rightarrow z_k} \frac{1}{\left(n-1\right)!} \frac{\mathrm{d}^{n-1}}{\mathrm{d}z^{n-1}} \left[f(z) (z-z_k) \right]
	\end{equation}
For first order poles (the cases considered below), $n=1$, so that 
	\begin{equation}
		\oint \mathrm{d}z f(z) = 2\pi i \sum_k \lim_{z \rightarrow z_k} f(z) (z-z_k)
	\end{equation}
\end{subequations}
Note that in \autoref{eq:generalcontour}, the integrals are closed along a contour which wraps all the poles.
